Exploring dimensions of biodiversity in Japanese ferns

### 2021-05-28 ms/manuscript.Rmd
Save or write plots depending on Rmd output format

```{r japan-map-save}
# Plot map of Japan with labels of relevant places

@@

# Make map
japan_map_plot <- ggplot(japan) + 
  geom_sf() +
  geom_richtext(
    data = japan_points, 
    aes(x = lon_adj, y = lat_adj, label = label),
    size = 2.5,
    fill = NA, label.color = NA, # remove background and outline
    label.padding = grid::unit(rep(0, 4), "pt") # remove padding)
  ) +
  geom_point(
    data = filter(japan_points, query == "Tokyo"),
    aes(x = lon, y = lat)
  ) +
  map_theme()

ggsave(result_file(figure("japan-map"), "pdf"), japan_map_plot, width = 5, height = 5, units = "in")
```

@@

```{r raw-diversity-save}
# Richness
a <- ggplot() +
  geom_sf(data = japan, fill = "grey90", color = "black", lwd = 0.25) +
  geom_sf(data = biodiv_ferns_spatial, aes(fill = richness), color = "transparent", alpha = 0.8) +
  annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**A** Richness</span>", hjust = "inward", vjust = "inward", color = "white") + 
  labs(fill = "No. taxa") +
  scale_fill_scico(palette = "lajolla") + 
  scale_x_continuous(breaks = c(125, 135, 145)) +
  scale_y_continuous(breaks = c(25, 35, 45)) +
  map_theme()

@@

raw_diversity_plot <- a + b + c +
  plot_layout(nrow = 1)

ggsave(result_file(figure("raw-diversity"), "pdf"), width = 7, height = 4, units = "in", raw_diversity_plot)
```

@@

```{r rand-diversity-save}
# PD
a <- ggplot() +
  geom_sf(data = japan, fill = "grey90", color = "black", lwd = 0.25) +
  geom_sf(data = biodiv_ferns_spatial, aes(fill = pd_signif), color = "transparent", alpha = 0.8) +
  annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**A** PD</span>", hjust = "inward", vjust = "inward", color = "white") + 
  scale_fill_manual(values = signif_cols) +
  map_theme() +
  theme(
    axis.text.x = element_blank(),
    legend.title = element_blank()
    )

@@

rand_diversity_plot <- a + b + c + d + 
  plot_layout(ncol = 2, nrow = 2, guides = 'collect') &
  theme(legend.position = "bottom")

ggsave(result_file(figure("rand-diversity"), "pdf"), rand_diversity_plot, width = 7, height = 8, units = "in")
```

@@

```{r endemism-save, fig.width = 5, fig.height = 5}
endemism_plot <- ggplot() +
  geom_sf(data = japan, fill = "grey90", color = "black", lwd = 0.25) +
  geom_sf(data = biodiv_ferns_spatial, aes(fill = endem_type), color = "transparent", alpha = 0.8) +
  scale_fill_manual(values = canape_cols) +
  map_theme() +
  theme(legend.title = element_blank())

ggsave(result_file(figure("endemism"), "pdf"), endemism_plot, width = 5, height = 5, units = "in")
```

@@

```{r bioregions-save}
# Taxonomic bioregions
a <- ggplot() +
  geom_sf(data = japan, fill = "grey90", color = "black", lwd = 0.25) +
  geom_sf(data = biodiv_ferns_spatial, aes(fill = taxonomic_cluster), color = "transparent", alpha = 0.8) +
  annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**A** Taxonomic</span>", hjust = "inward", vjust = "inward", color = "white") + 
  labs(fill = "Cluster") +
  scale_fill_manual(values = bioregion_cols) +
  map_theme()

@@

bioregions_plot <- a + b +
  plot_layout(ncol = 2)

ggsave(result_file(figure("bioregions"), "pdf"), bioregions_plot, width = 7, height = 5, units = "in")
```

@@

```{r div-by-tax-region-save}
data <-
biodiv_ferns_spatial %>%
  filter(taxonomic_cluster %in% c(1,2,3,4)) %>%
  as_tibble %>%
  select(pd_obs_z, rpd_obs_z, fd_obs_z, rfd_obs_z, pe_obs_p_upper, taxonomic_cluster)

@@

div_by_tax_region_plot <- a + b + c + d + e +
  plot_layout(ncol = 2) +
plot_annotation(
  tag_levels = 'A'
) &
  theme(legend.position = "none")

ggsave(result_file(figure("div-by-tax-region"), "pdf"), div_by_tax_region_plot, width = 7, height = 7, units = "in")
```

@@

```{r conserv-status-save, fig.width = 4, fig.height = 4}
# Plot protected areas
conserv_status_plot <- signif_cells_protected_area %>%
  mutate(
    metric = case_when(
      metric == "pd" ~ "PD",
      metric == "fd" ~ "FD",
      metric == "pe" ~ "PE",
      metric == "richness" ~ "Richness",
    ),
    status = str_to_sentence(status)) %>%
  mutate(metric = factor(metric, levels = c("PE", "FD", "PD", "Richness"))) %>%
  ggplot(aes(x = percent_protected, y = metric, fill = status)) +
  geom_col(position = position_dodge()) +
  geom_vline(xintercept = total_percent_protected_medium, color = "grey40", linetype = 3) +
  geom_vline(xintercept = total_percent_protected_high, color = "grey40", linetype = 2) +
  scale_fill_viridis_d() +
  scale_x_continuous(labels = scales::percent) +
  labs(
    x = "Percent protected",
    y = "Biodiversity metric",
    fill = "Protection status"
  ) +
  theme_bw() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank(),
    legend.position = "bottom"
  )

ggsave(result_file(figure("conserv-status"), "pdf"), conserv_status_plot, width = 4, height = 4, units = "in")
```

>>>>

```{r japan-map-make}
# Plot map of Japan with labels of relevant places

@@

# Make map
japan_map_plot <- ggplot(japan) + 
  geom_sf() +
  geom_richtext(
    data = japan_points, 
    aes(x = lon_adj, y = lat_adj, label = label),
    size = 2.5,
    fill = NA, label.color = NA, # remove background and outline
    label.padding = grid::unit(rep(0, 4), "pt") # remove padding)
  ) +
  geom_point(
    data = filter(japan_points, query == "Tokyo"),
    aes(x = lon, y = lat)
  ) +
  map_theme()
```

```{r japan-map-save, eval = params$doc_type == "doc"}
# Save plot if rendering to doc
ggsave(result_file(figure("japan-map"), "pdf"), japan_map_plot, width = 5, height = 5, units = "in")
```

@@

```{r raw-diversity-make}
# Richness
a <- ggplot() +
  geom_sf(data = japan, fill = "grey90", color = "black", lwd = 0.25) +
  geom_sf(data = biodiv_ferns_spatial, aes(fill = richness), color = "transparent", alpha = 0.8) +
  annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**A** Richness</span>", hjust = "inward", vjust = "inward", color = "white") + 
  labs(fill = "No. taxa") +
  scale_fill_scico(palette = "lajolla") + 
  scale_x_continuous(breaks = c(125, 135, 145)) +
  scale_y_continuous(breaks = c(25, 35, 45)) +
  map_theme()

@@

raw_diversity_plot <- a + b + c +
  plot_layout(nrow = 1)
```

```{r raw-diversity-save, eval = params$doc_type == "doc"}
# Save plot if rendering to doc
ggsave(result_file(figure("raw-diversity"), "pdf"), width = 7, height = 4, units = "in", raw_diversity_plot)
```

@@

```{r rand-diversity-make}
# PD
a <- ggplot() +
  geom_sf(data = japan, fill = "grey90", color = "black", lwd = 0.25) +
  geom_sf(data = biodiv_ferns_spatial, aes(fill = pd_signif), color = "transparent", alpha = 0.8) +
  annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**A** PD</span>", hjust = "inward", vjust = "inward", color = "white") + 
  scale_fill_manual(values = signif_cols) +
  map_theme() +
  theme(
    axis.text.x = element_blank(),
    legend.title = element_blank()
  )

@@

rand_diversity_plot <- a + b + c + d + 
  plot_layout(ncol = 2, nrow = 2, guides = 'collect') &
  theme(legend.position = "bottom")
```

```{r rand-diversity-save, eval = params$doc_type == "doc"}
# Save plot if rendering to doc
ggsave(result_file(figure("rand-diversity"), "pdf"), rand_diversity_plot, width = 7, height = 8, units = "in")
```

@@

```{r endemism-make}
endemism_plot <- ggplot() +
  geom_sf(data = japan, fill = "grey90", color = "black", lwd = 0.25) +
  geom_sf(data = biodiv_ferns_spatial, aes(fill = endem_type), color = "transparent", alpha = 0.8) +
  scale_fill_manual(values = canape_cols) +
  map_theme() +
  theme(legend.title = element_blank())
```

```{r endemism-save, eval = params$doc_type == "doc"}
# Save plot if rendering to doc
ggsave(result_file(figure("endemism"), "pdf"), endemism_plot, width = 5, height = 5, units = "in")
```

@@

```{r bioregions-make}
# Taxonomic bioregions
a <- ggplot() +
  geom_sf(data = japan, fill = "grey90", color = "black", lwd = 0.25) +
  geom_sf(data = biodiv_ferns_spatial, aes(fill = taxonomic_cluster), color = "transparent", alpha = 0.8) +
  annotate(geom="richtext", y = Inf, x = -Inf, label = "<span style='color:black'>**A** Taxonomic</span>", hjust = "inward", vjust = "inward", color = "white") + 
  labs(fill = "Cluster") +
  scale_fill_manual(values = bioregion_cols) +
  map_theme()

@@

bioregions_plot <- a + b +
  plot_layout(ncol = 2)
```

```{r bioregions-save, eval = params$doc_type == "doc"}
# Save plot if rendering to doc
ggsave(result_file(figure("bioregions"), "pdf"), bioregions_plot, width = 7, height = 5, units = "in")
```

@@

```{r div-by-tax-region-make}
data <-
  biodiv_ferns_spatial %>%
  filter(taxonomic_cluster %in% c(1,2,3,4)) %>%
  as_tibble %>%
  select(pd_obs_z, rpd_obs_z, fd_obs_z, rfd_obs_z, pe_obs_p_upper, taxonomic_cluster)

@@

div_by_tax_region_plot <- a + b + c + d + e +
  plot_layout(ncol = 2) +
  plot_annotation(
    tag_levels = 'A'
  ) &
  theme(legend.position = "none")
```

```{r div-by-tax-region-save, eval = params$doc_type == "doc"}
# Save plot if rendering to doc
ggsave(result_file(figure("div-by-tax-region"), "pdf"), div_by_tax_region_plot, width = 7, height = 7, units = "in")
```

@@

```{r conserv-status-make}
# Plot protected areas
conserv_status_plot <- signif_cells_protected_area %>%
  mutate(
    metric = case_when(
      metric == "pd" ~ "PD",
      metric == "fd" ~ "FD",
      metric == "pe" ~ "PE",
      metric == "richness" ~ "Richness",
    ),
    status = str_to_sentence(status)) %>%
  mutate(metric = factor(metric, levels = c("PE", "FD", "PD", "Richness"))) %>%
  ggplot(aes(x = percent_protected, y = metric, fill = status)) +
  geom_col(position = position_dodge()) +
  geom_vline(xintercept = total_percent_protected_medium, color = "grey40", linetype = 3) +
  geom_vline(xintercept = total_percent_protected_high, color = "grey40", linetype = 2) +
  scale_fill_viridis_d() +
  scale_x_continuous(labels = scales::percent) +
  labs(
    x = "Percent protected",
    y = "Biodiversity metric",
    fill = "Protection status"
  ) +
  theme_bw() +
  theme(
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank(),
    legend.position = "bottom"
  )
```

```{r conserv-status-save, eval = params$doc_type == "doc"}
# Save plot if rendering to doc
ggsave(result_file(figure("conserv-status"), "pdf"), conserv_status_plot, width = 4, height = 4, units = "in")
```
